Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2021-01-12 Nicolas.tex
Greg changes, diff to see them

\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
    unknown, and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of the SN~Ia population
    evolution. In this paper, we study the dependence with redshift of the SN~Ia
    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
    probe its potential redshift drift. The SN stretch has been shown to
    strongly correlate with the SN environment, notably with stellar age
    tracers. We model the underlying stretch distribution as a function of
    redshift, using the evolution of the fraction of young and old SNe~Ia as
    predicted by \cite{rigault2020}, and assuming constant underlying stretch
    distribution for each age population \textbf{consisting} of Gaussian
    mixtures. We test our prediction against published samples \textbf{cut} to
    have marginal magnitude selection effects, so that any observed change is
    indeed of astrophysical and not observational origin. \textbf{In this first
    study, there are indications} that the underlying SN~Ia stretch distribution
    is evolving as a function of redshift, and that the young/old drifting model
    is a better description of the data than any time-constant model, including
    the sample-based asymmetric distributions usually used to correct Malmquist
    bias. The favored underlying stretch model is the bimodal one derived from
    \cite{rigault2020}, \textbf{composed of} a high-stretch mode shared by both
    young and old environments, and a low-stretch mode exclusive to old
    environments. The precise impact of the redshift evolution of the SN Ia
    population intrinsic properties on cosmology remains to be studied. Yet, the
    astrophysical drift of the SN stretch distribution does affect current
    Malmquist bias corrections and thereby distances derived from SN affected by
selection effects. We highlight that such a bias will increase with surveys
covering increasingly larger redshift ranges, which is particularly important
for LSST.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate ($z<0.5$), when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant ($H_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES project,][]{riess2009,
riess2016} is $\sim5\sigma$ higher than what is predicted from cosmic microwave
background (CMB) data measured by Planck assuming the standard $\L$CDM
\citep{planck2018, riess2019, reid2019}, or when the SN luminosity is anchored
at intermediate redshift by the baryon acoustic oscillation (BAO) scale
\citep{feeney2019}. While using the tip of the red giant branch technique in
place of the Cepheids seem to favor an intermediate value of $H_0$
\citep{freedman2019, freedman2020}, time delay measurements from strong lensing
seem to also favor high $H_0$ values \citep{wong2019}.

The $H_0$ tension has received a lot of attention, as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this $H_0$
tension when accounting for all other probes \citep{knox2019}, but see e.g.
\cite{poulin2019}. Alternatively, systematic effects affecting one or several
of the aforementioned analyses could also explain the tension.
\cite{rigault2015} suggested that SNe Ia from the Cepheid calibrator sample
differ by construction from the Hubble flow sample ones as the former
strongly favors young stellar populations, while the latter not. This
selection effect would impact the derivation of $H_0$ if SNe~Ia from young and
older environments differ in average standardized magnitudes.

For the last decades, numerous analyses have studied the relation between SNe~Ia
and host properties, finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar mass, SNe~Ia from high-mass host being
brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
betoule2014, rigault2020, kim19}. This mass-step correction is currently used in
cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including for
deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
between the SNe and their host remains unclear when using global properties such
as the host stellar mass, which raises the question of the accuracy of the
correction. More recently, studies have used the local SN environment to probe
more direct connections between the SN and its environments \citep{rigault2013},
showing that local age tracers such as the Local specific Star Formation Rate
(LsSFR) or the local color are more strongly correlated with the standardized SN
magnitude \citep{rigault2020, roman2018, kim18}, suggesting age as the driving
parameter underlying the mass-step. If true, this would have a significant
impact for cosmology, since the environmental correction to apply to SN
standardization could strongly vary with redshift. \citep{rigault2013,
childress2014, scolnic2018a}. Yet, the importance of local SN environmental
studies remains highly debated \cite[e.g.][]{jones2015, jones2019} and
especially the impact of such an astrophysical bias has on the derivation of
$H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.

The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
the relative SNe~Ia rate in galaxies could be explained if two populations
existed, one young, following the host star formation activity, and one old
following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
model). In \cite{rigault2020} we used the LsSFR to classify which are the
younger (those with a high LsSFR) and which are the older (those with low
LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN stretch has
been known to be strongly correlated with the SN host properties
\citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed
since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
the ``A+B'' model and the connection between SN stretch and host properties,
\cite{howell2007} first discussed the potential redshift drift of the SN stretch
distribution. In this paper we revisit this analysis with the most recent SNe~Ia
dataset.

In this paper, we take a step aside to probe the validity of our modeling of the
SN population, which we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2020}: one old and one younger, the former
having on average lower lightcurve stretches and being brighter after
standardization. We use the correlation between the SN age, as probed by the
LsSFR, and the SN stretch to model the expected evolution of the underlying SN
stretch distribution as a function of redshift. This modeling relies on three
assumptions: (1) there are two distinct populations of SNe~Ia; (2) the relative
fraction of each of these populations as a function of redshift follows the
model presented in \cite{rigault2020} and (3) the underlying distribution of
stretch for each age sample is constant. This paper aims at testing this
specific model with datasets from the literature. Note that the progenitor age
as traced by the LsSFR seems to capture physical feature intrinsic to the
progenitor and/or explosion mechanism that the stretch alone is not capturing
\citep{nordin2018}.

@@

\section{Complete Sample Construction}\label{sec:sample}
\subsection{Applying redshift cuts}\label{ssec:cuts}

We base our analysis on the most recent comprehensive SNe~Ia compilation, the
Pantheon catalog from \cite{scolnic2018a}. A naive approach to test the SN
stretch redshift drift would be to simply compare the observed SN stretch
distributions in a few bins of redshift. In practice, however, differential
selection effects are affecting the observed SN stretch distributions. Indeed,
because the observed SN~Ia magnitude correlates with the lightcurve stretch (and
color), the first SNe~Ia that a magnitude-limited survey will miss are the
lowest-stretch (and reddest) ones. Consequently, if magnitude-related selection
effects are not accounted for, one might confuse true population drift with
survey properties, and conversely.

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring typing
and host redshift, the selection effects of magnitude-limited surveys should be
negligible below a given redshift at which even the faintest SNe~Ia can be
observed. \textbf{In practice such spectroscopic following has not always been
the norm as discussed below.} In contrast, targeted surveys have highly complex
selection functions and will be discarded from our analysis. Modern SN cosmology
samples such as the Pantheon one are now dominated by magnitude-limited surveys
\textbf{which have easier to model selection functions}.

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
\citet{campbell2013} for similar contours, the second using a more conservative
$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is
$M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
along the aforementioned ellipse given the standardization coefficient
$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is
that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one ought to detect
this object typically 5 days before and a week after peak to build a suitable
lightcurve, the effective limiting standardized absolute magnitude is
approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit
$m_{\lim}$ of a magnitude limited survey, one can derive the maximum redshift
$z_{\lim}$ above which the faintest SNe~Ia will be missed using the relation
between apparent magnitude, redshift and absolute magnitude $\mu(z_{\lim}) =
m_{\lim} - M_{\lim}$.

@@

\textbf{We will thus consider a set of cuts that will define a first fiducial
    sample, taking the limits as initially suggested by the previous
    completeness analysis. \textbf{However,} as this solution might be over
    simplistic to create a complete sample, e.g., for it ignores follow-up
    efficiency, we also consider another set of cuts to define a so-called
    conservative sample. This latter is smaller and therefore less statistically
    constraining, but also inherently less prone to selection effects. If the
    redshift drift is still significant in the conservative sample, it would be
    even more meaningful in a carefully-tailored selection-free sample. The goal
    of this study is to build the first implementation of a SN Ia properties
    drift modeling before refining it on a more complex, yet more accurate,
follow-up analysis also accounting for precise selection modeling using, e.g.,
the SNANA package \citep{SNANA}.}

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, that has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
\citet[][see their Section~5]{perrett2010} suggests however a lower limit of
$z_{\lim}=0.55$. \textbf{We will therefore use $z=0.60$ and $z=0.55$ as redshift
limits for the fiducial and conservative cuts, respectively.}

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. \textbf{If we were to be
    conservative, this figure would also suggest of a more stringent
$z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
conservative cuts, respectively.}

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
SDSS surveys \textbf{had to contend with limited spectroscopic resources. As
discussed in} \citet[][Section~2]{kessler2009}, during the first year of SDSS,
SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
these analyses, \textbf{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
fiducial and conservative cuts, respectively.}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from the
        SDSS, PS1 and SNLS dataset respectively (data from Pantheon,
        \citealt{scolnic2018a}). The colored parts represent the distribution of
        SNe~Ia kept in our analysis for they are supposedly free from selection
    bias (see Section~\ref{sec:sample}). The darker (resp. lighter) color
responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

@@

Finally, we include the HST sample from Pantheon \citep{strolger04}.
\textbf{These high-redshift SNe are of great interest as they provide the
    greatest leverage for testing evolution. While at these redshifts the
    supernovae typing is challenging, the target classification was robust
    enough to include them within the cosmological analysis \citep{scolnic2018a}
    and we do not add further quality cuts. Section~\ref{sec:results} highlights
    that, while compatible with it, our results are not dependent on the
    inclusion of this dataset.}

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe
that the fraction of low-stretch SNe (typically $x_1 < -1$) seems to decrease as
a function of redshift; this is confirmed in Fig.~\ref{fig:modelall} in which}
the evolution of the mean stretch is shown, \textbf{with} the data split in
redshift bins of regular sample size. We see that SNe~Ia at higher redshift have
on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
stretch distribution is \textbf{evolving with redshift}.

@@

\textbf{In section~\ref{ssec:cuts}, we have built volume-limited samples from a
    set of magnitude-limited ones, using simple redshift cuts. This simplistic
    approach is statistically sub-optimal, but should be enough to demonstrate
    our key finding: a redshift evolution of stretch compatible with
    \cite{rigault2020} model. However, the possibility remains that a complex
    selection function related to follow-up efficiencies below our fiducial (or
    even conservative) cuts could still affect our sample, making it not fully
volume-limited; this would, in turn, bias our conclusion on the astrophysical
drift of the SNe~Ia population. We now look at this possibility.}

\textbf{To test for the existence of potential leftover selection biases in our
    sample, we compare the stretch and color distributions of the SNe~Ia
    originating from different datasets over overlapping redshift ranges: these
    distributions should be similar. Note that the redshift range has to be
    narrow enough so that any drift would be negligible.}

\textbf{The two samples that overlap the most in redshift are PS1 and SDSS in
    the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
    corresponds to the 146 SNe~Ia from SDSS at the high redshift end and thus
    the most likely to be affected by residual selection effects (see the
    corresponding discussion in section~\ref{ssec:cuts}). Over that same
    redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
    thus unlikely to have any selection issue. To identify potential
    inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib}
    (upper panels) compares the stretch and color distribution of both these
    surveys. The associated Kolmogorov-Smirnov (KS) similarity test $p$-values
are too high ($>10\%$) to conclude on any inconsistency, in agreement with the
visual impression from Fig.~\ref{fig:distrib}.}

\textbf{We perform a similar analysis for PS1 and SNLS over the redshift range
    $0.20 < z < 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same
    conclusion can be drawn: there is no substantial sign of discrepancy in the
    stretch and color distributions between the low- and high-end of the SNLS
    and PS1 datasets, respectively. Nonetheless, the small size of the SNLS
    dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of
    the test, and only strong deviation would be noticeable. Extending the
redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
allows to increase the SNLS subsample to 31, yet the stretch $p$-values remains
high (34\%) showing no sign of inconsistency.}

@@

\textbf{We finally highlight that the SNe~Ia color is more prone to selection
    effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
    Fig.~3 of \cite{kessler2017}. Hence, since the comparison of color
    distributions shows no significant hint of leftover selection effect, it
    further supports our claim that our simple redshift-based selection criteria
    is sufficient to build a complete SNe~Ia sample required to test the
    redshift evolution of the stretch distribution.}

@@

\bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G.,
Dahlen, T., et al.\ 2004, \apj, 613, 200

>>>>

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of SN~Ia population
    evolution. In this paper, we study the dependence with redshift of the SN~Ia
    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
    probe its potential redshift drift. The SN stretch has been shown to
    strongly correlate with the SN environment, notably with stellar age
    tracers. We model the underlying stretch distribution as a function of
    redshift, using the evolution of the fraction of young and old SNe~Ia as
    predicted by \cite{rigault2020}, and assuming constant underlying stretch
    distribution for each age population \textbf{consisting} of Gaussian
    mixtures. We test our prediction against published samples \textbf{cut} to
    have marginal magnitude selection effects, so that any observed change is
    indeed of astrophysical and not observational in origin. \textbf{In this first
    study, there are indications} that the underlying SN~Ia stretch distribution
    is evolving as a function of redshift, and that the young/old drifting model
    is a better description of the data than any time-constant model, including
    the sample-based asymmetric distributions often used to correct Malmquist
    bias. The favored underlying stretch model is the bimodal one derived from
    \cite{rigault2020}, \textbf{composed of} a high-stretch mode shared by both
    young and old environments, and a low-stretch mode exclusive to old
    environments. The precise impact of the redshift evolution of the SN Ia
    population intrinsic properties on cosmology remains to be studied. Yet, the
    astrophysical drift of the SN stretch distribution does affect current
    Malmquist bias corrections and thereby distances derived from SN affected by
observational selection effects. We highlight that such a bias will increase with surveys
covering increasingly larger redshift ranges, which is particularly important
for LSST.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate ($z<0.5$), when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant ($H_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ derived when
the SNe~Ia are anchored to Cepheids \citep[the SH0ES project,][]{riess2009,
riess2016} is $\sim5\sigma$ higher than what is predicted from cosmic microwave
background (CMB) data measured by Planck assuming the standard $\L$CDM
\citep{planck2018, riess2019, reid2019}, or when the SN luminosity is anchored
at intermediate redshift by the baryon acoustic oscillation (BAO) scale
\citep{feeney2019}. While using the tip of the red giant branch technique in
place of the Cepheids seem to favor an intermediate value of $H_0$
\citep{freedman2019, freedman2020}, time delay measurements from strong lensing
seem to also favor high $H_0$ values \citep{wong2019}.

The $H_0$ tension has received a lot of attention, as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this $H_0$
tension when accounting for all other probes \citep{knox2019}, but see e.g.
\cite{poulin2019}. Alternatively, systematic effects affecting one or several
of the aforementioned analyses might explain at least some of this tension.
\cite{rigault2015} suggested that SNe Ia from the Cepheid calibrator sample
differ by construction from those in the Hubble flow sample, as the former
strongly favors young stellar populations, while the latter does not. This
selection effect would impact the derivation of $H_0$ if SNe~Ia from young and
older environments differ in average standardized magnitudes.

%For the last decades, numerous analyses have studied 
The relation between SNe~Ia
and their host galaxy properties has been studied extensively. The first key finding
was that that the standardized SNe~Ia magnitudes
significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass host 
galaxies being
brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
betoule2014, rigault2020, kim19}. This mass-step correction is currently used in
cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including for
deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
between the SNe and their host galaxies remains unclear, particularly when using global properties such
as the host stellar mass. This in turn raises the question of the accuracy of such
corrections since host galaxy global properties evolve with redshift. 
More recently, studies have used the local SN host galaxy environment to probe
more direct connections between the SNe and their host galaxy environments \citep{rigault2013},
showing that local age tracers such as the Local specific Star Formation Rate
(LsSFR) or the local color are more strongly correlated with the standardized SN
magnitude \citep{rigault2020, roman2018, kim18}. Further reinforcing this
connection is the identification of SN~Ia spectral features that are 
correlated with LsSFR \citep{nordin2018}. These results suggest age as the driving
parameter underlying the mass-step. If true, this would have a significant
impact for cosmology, since the environmental correction to apply to SN
standardization could strongly vary with redshift. \citep{rigault2013,
childress2014, scolnic2018a}. Yet, the importance of local SN environmental
studies remains highly debated \cite[e.g.][]{jones2015, jones2019} and
especially the impact of such an astrophysical bias has on the derivation of
$H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.

The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
the relative SNe~Ia rate in galaxies could be explained if two populations
existed, one young, following the host star formation activity, and one old
following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
model). In \cite{rigault2020} we used the LsSFR to classify which are the
younger (those with a high LsSFR) and which are the older (those with low
LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN stretch has
been known to be strongly correlated with the SN host galaxy properties
\citep{hamuy1996, hamuy2000}. This correlation that has been extensively confirmed
since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
the ``A+B'' model and the connection between SN stretch and host galaxy properties,
\cite{howell2007} first discussed the potential redshift drift of the SN stretch
distribution. In this paper we revisit this question using the most recent SNe~Ia
datasets.

In this paper, we take a step aside to probe the validity of our modeling of the
SN population, which we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2020}: one old and one younger, the former
having on average lower lightcurve stretches and being brighter after
standardization. We use the correlation between the SN age, as probed by the
LsSFR, and the SN stretch to model the expected evolution of the underlying SN
stretch distribution as a function of redshift. This modeling relies on three
assumptions: (1) there are two distinct populations of SNe~Ia; (2) the relative
fraction of each of these populations as a function of redshift follows the
model presented in \cite{rigault2020} and (3) the underlying distribution of
stretch for each age sample is constant. This paper aims at testing this
specific model with datasets from the literature. 
% Note that the progenitor age
% as traced by the LsSFR seems to capture physical feature intrinsic to the
%progenitor and/or explosion mechanism that the stretch alone is not capturing
% \citep{nordin2018}.

@@

\section{Complete Sample Construction}\label{sec:sample}

The ideal SN~Ia sample for studying this question would be a very deep, large-area,
volume-limited sample. This would capture the true underlying stretch distribution
function, and we would then study how it evolves with redshift. No such sample exists, 
so first we must construct subsamples that are as near to volume-limited as possible 
from existing high-redshift SN~Ia samples.

\subsection{Applying redshift cuts}\label{ssec:cuts}

We base our analysis on the most recent comprehensive SNe~Ia compilation, the
Pantheon catalog from \cite{scolnic2018a}. A naive approach to test the SN
stretch redshift drift would be to simply compare the observed SN stretch
distributions in a few bins of redshift. In practice, however, differential
observational selection effects will affect the observed SN stretch distributions. Indeed,
because the observed SN~Ia magnitude correlates with the lightcurve stretch (and
color), the first SNe~Ia that a magnitude-limited survey will miss are the
lowest-stretch (and reddest) ones. Consequently, if magnitude-related observational selection
effects are not accounted for, one might confuse true population drift with
survey properties, and conversely.

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring type SN types
and host galaxy redshifts, the observation selection effects of magnitude-limited surveys should be
negligible below a given redshift at which even the faintest normal SNe~Ia can be
observed. \textbf{Aiding in the construction of nearly volume-limited 
subsamples is the fact that the SN~Ia population trails off towards fainter SNe~Ia.
A complication is that complete spectroscopic follow-up has not always been
the norm, as discussed below.}
In contrast, targeted surveys have highly complex observational
selection functions and so are discarded from our analysis. High-redshift SN cosmology
samples, such as Pantheon, are predominately from magnitude-limited surveys from
which volume-limited SN~Ia subsamples can be constructed.
%\textbf{which have easier to model observational selection functions}.

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
parent distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
\citet{campbell2013} for similar contours, the second using a more conservative
$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is
$M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
along the aforementioned ellipse given the standardization coefficient
$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is
that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one ought to detect
this object typically 5 days before and a week after peak to build a suitable
lightcurve, the effective limiting standardized absolute magnitude is
approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit
$m_{\lim}$ of a magnitude limited survey, one can derive the maximum redshift
$z_{\lim}$ above which the faintest SNe~Ia will be missed using the relation
between apparent magnitude, redshift and absolute magnitude $\mu(z_{\lim}) =
m_{\lim} - M_{\lim}$.

@@

\textbf{We will thus consider a set of cuts that will define a first fiducial
    sample, taking the limits as initially suggested by the previous
    completeness analysis. \textbf{However,} as this solution might be an overly
    simplified way to create a complete sample, e.g., because it ignores spectroscopic follow-up
    in efficiency for redshifts below $z_{lim}$, we also consider another set of cuts to define a so-called
    ``conservative'' sample. The latter is smaller and therefore less statistically
    constraining, but also even less prone to observational selection effects. If the
    redshift drift is still significant in the conservative sample, it would be
    even more meaningful in a carefully-tailored selection-free sample. These
    samples are adequate for the goal
    of this study, which is to develop a first implementation of a model for drift in
    SN Ia properties. If fruitful, the sample selection can later be refined, if needed, with a
    more detailed
    %drift modeling before refining it on a more complex, yet more accurate,
%follow-up analysis also accounting for precise 
model of the observational selection, e.g., using
the SNANA package \citep{SNANA}.}

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, which has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
\citet[][see their Section~5]{perrett2010} suggests however a lower limit of
$z_{\lim}=0.55$. \textbf{We will therefore use $z=0.60$ and $z=0.55$ as redshift
limits for the fiducial and conservative samples, respectively, for SNLS.}

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. \textbf{If we were to be
    conservative, this figure would also suggest of a more stringent
$z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
conservative samples, respectively, for PS1.}

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
SDSS surveys \textbf{had to contend with limited spectroscopic resources. As
discussed in} \citet[][Section~2]{kessler2009}, during the first year of SDSS,
SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Based on this,
\textbf{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
fiducial and conservative samples, respectively, for SDSS.}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from the
        SDSS, PS1 and SNLS dataset respectively (data from Pantheon,
        \citealt{scolnic2018a}). The colored parts represent the distribution of
        SNe~Ia kept in our analysis for they are supposedly free from observational selection
    bias (see Section~\ref{sec:sample}). The darker (resp. lighter) color
responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

@@

Finally, we include the HST sample from Pantheon \citep{strolger04}.
\textbf{These high-redshift SNe are of great interest as they provide the
    greatest leverage for testing evolution. While at these redshifts the
    supernovae typing is challenging, the target classification was robust
    enough to include them within the cosmological analysis \citep{scolnic2018a}
    and we do not impose further cuts. Section~\ref{sec:results} highlights
    that, while compatible with it, our results are not dependent on the
    inclusion of this dataset.}

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe here
that the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as
a function of redshift; this is confirmed in Fig.~\ref{fig:modelall} in which}
the evolution of the mean stretch is shown, \textbf{with} the data split in
redshift bins of regular sample size. We see that SNe~Ia at higher redshift have
on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
stretch distribution is \textbf{evolving with redshift}.

@@

\textbf{In section~\ref{ssec:cuts}, we have built volume-limited samples from a
    set of magnitude-limited ones, using simple redshift cuts. This simplified
    approach is statistically sub-optimal, but should be enough to test
    our key question: whether redshift evolution of stretch is compatible with the
    \cite{rigault2020} model. However, the possibility remains that a complex
    observational selection function related to spectroscopic follow-up efficiencies below our fiducial (or
    even conservative) redshift cuts could still affect our sample, making it not fully
volume-limited; this would, in turn, bias our conclusion on the astrophysical
drift of the SNe~Ia population. We now look at this possibility.}

\textbf{To test for the existence of potential leftover observational selection biases in our
    sample, we compare the stretch and color distributions of the SNe~Ia
    originating from different datasets having overlapping redshift ranges: these
    distributions should be similar if they reflect the underlying parent population. 
    Note that the redshift range has to be
    narrow enough so that any drift would be negligible.}

\textbf{The two samples that overlap the most in redshift are PS1 and SDSS in
    the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping subsample
    consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
    the most likely to be affected by residual observational selection effects (see the
    corresponding discussion in section~\ref{ssec:cuts}). Over that same
    redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
    thus unlikely to have any observational selection issue. To identify potential
    inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib}
    (upper panels) compares the stretch and color distribution of both these
    surveys. The resulting Kolmogorov-Smirnov (KS) similarity test $p$-values
($p >10\%$) do not support any inconsistency, in agreement with the
visual impression from Fig.~\ref{fig:distrib}.}

\textbf{We perform a similar analysis for PS1 and SNLS over the redshift range
    $0.20 < z < 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same
    conclusion can be drawn: there is no substantial sign of discrepancy in the
    stretch and color distributions between the low- and high-end of our fiducial SNLS
    and PS1 samples, respectively. Nonetheless, the small size of the SNLS
    dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of
    this test, and only a strong deviation would be noticeable. Extending the
redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
allows to increase the SNLS subsample to 31, yet the stretch $p$-values remains
high (34\%) showing no sign of inconsistency.}

@@

\textbf{We finally highlight that the SNe~Ia color is more prone to observational selection
    effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
    Fig.~3 of \cite{kessler2017}. Hence, since the comparison of color
    distributions shows no significant hint of leftover observational selection effect, it
    further supports our claim that our simple redshift-based selection criteria
    are sufficient to build the complete SNe~Ia samples required to test the
    redshift evolution of the stretch distribution.}

@@

\bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G.,
